### 2015-12-11 paper/4-performance.tex
updated interconnect parallel graph

In this section, we evaluate performance of dxex using its different synchronization protocols.
We show that the inclusion of multiple synchronization protocols does not decrease performance significantly.
To this end, we compare to adevs, currently one of the most efficient simulation kernels~\cite{DEVSSurvey}.
CPU time and memory usage is compared for both sequential and parallel simulation.

We start off with a comparison of sequential simulation, to show how adevs and dxex relate in this simple case.
For the parallel simulation benchmarks, results are presented for both conservative and optimistic synchronization.

For all benchmarks, results are well within a 1\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3 -flto}'').
To guarantee comparable results, no I/O was performed during the benchmarking phase such that computation was CPU-bound.
Simulation traces are used to verify that both adevs and dxex models have exactly the same behaviour.
Benchmarks were performed using Linux, but our simulation tool works equally well on Windows and Mac.

\subsection{Benchmarks}
We use a selection of benchmarks, based on those found in the literature.
Three different types of benchmark are defined:
\begin{enumerate}
    \item The \textit{Queue} model, based on the HI model of DEVStone~\cite{DEVStone}, creates a chain of hierarchically nested atomic \textsf{DEVS} models.
          A single generator will push events into the queue, which get processed by the processors after a fixed or random delay.
          It takes two parameters: width and depth, which determine the width and depth of the hierarchy.
          This benchmark shows how the complexity of the simulation kernel behaves for an increasing amount of atomic models, and an increasingly deep hierarchy.
          An example for width 2 and depth 3 is shown in Figure~\ref{fig:queue_model}.

    \item The \textit{PHOLD} model, presented by~\cite{PHOLD}, creates $n$ atomic models, where each model has exactly $n-1$ output ports.
          Couplings are made to directly connect each atomic model to every other atomic model.
          After a random delay, atomic models will send out an event to a randomly selected output port.
          Output port selection happens in two phases: first it is decided whether the event should be sent to an atomic model inside or outside of this coupled model.
          Afterwards, a uniform selection is made between the remaining ports.
          The model takes one parameter: the percentage of remote events, which determines the fraction of messages routed to other coupled models.
          This benchmark shows how the simulation kernel behaves in the presence of many local or remote events.
          An example for four models, split over two nodes, is shown in Figure~\ref{fig:PHOLD_model}.

    \item The \textit{HighInterconnect} model, a merge of the HI model of DEVStone~\cite{DEVStone} and PHOLD~\cite{PHOLD}, creates $n$ atomic models, where each model has exactly one output port.
          All models are still connected to one another, but all over the same port: every model will receive a generated event.
          The model takes one parameter: the number of models.
          This benchmark investigates the complexity of the event routing algorithm, and how the simulation kernel handles many simultaneous events.
          An example for four models is shown in Figure~\ref{fig:interconnect_model}.
\end{enumerate}

We opted to deviate from the DEVStone benchmark, as DEVStone tends towards unrealistic models since most transitions occur simultaneously.
In all of our benchmark models, simultaneous internal transitions are avoided as much as possible through the use of a random time advance.
DEVStone also uses an artificial load function, which easily skews the result: the simulation kernel overhead is dominated by the artificial load.

@@

\subsection{Sequential Simulation Execution Time}
Despite our core contribution being mainly in the parallel simulation, we still value a comprehensive comparison of sequential simulation results.
First, and foremost, as parallel simulation results are tightly linked to the sequential simulation results: parallel simulation merely adds a synchronization layer over multiple, essentially sequential, simulation kernels.
Second, parallel simulation results are validated through the use of adevs.
To provide a more comprehensive comparison to adevs in the parallel simulation benchmarks, sequential simulation results need to be compared.
Only the Queue and HighInterconnect models are relevant for sequential simulation, so we will not touch upon the PHold model yet.

\subsubsection{Queue}
In the Queue model, we increase both the width and depth simultaneously, causing a quadratic growth in the number of atomic models.
As can be seen in Figure~\ref{fig:Queue_benchmark}, dxex considerably outperforms adevs.
Through careful analysis of profiling results, we determined that adevs spends much time while handling simulation messages, while this is mostly avoided in dxex's differently designed simulation algorithm.
Both simulation tools have similar complexities, though dxex is much faster thanks to its more efficient simulation control algorithms.

@@

\subsubsection{HighInterconnect}
In the HighInterconnect model, we increase the number of atomic models, thus quadratically increasing the number of couplings.
As can be seen in Figure~\ref{fig:Interconnect_benchmark}, adevs now outperforms dxex by a fair margin.
Analysis showed that this slowdown is caused by the high amount of exchanged events: event creation is found to be much slower in dxex than it is in adevs, despite the use of memory pools in dxex.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{fig/queue_parallel.eps}
	\caption{Queue benchmark results for parallel simulation using 4 cores.}
	\label{fig:queue_benchmark_parallel}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{fig/phold_remotes.eps}
    \caption{Phold benchmark results for parallel simulation using 4 cores, with varying fraction of remote events.}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{fig/phold_voorlopig.eps}
    \caption{Phold benchmark results for parallel simulation using 4 cores, with varying amount of high-priority events.}
    \label{fig:phold_priority}
\end{figure}

>>>>

In this section, we evaluate the performance of our simulation tool.
We show that the inclusion of multiple synchronization protocols does not decrease our performance to the point where it is unusable.
To this end, we compare to adevs, one of the most efficient simulation kernels at this time~\cite{DEVSSurvey}.
A comparison is made for both the CPU and memory usage of both sequential simulation and parallel simulation.

We start off with a comparison of sequential simulation, to show how adevs and dxex relate in this simple case.
We not only show that our approach does not influence performance negatively, but also that our main simulation algorithm, similar to the one of PythonPDEVS, is significantly faster than the one found in adevs.
Similar differences, compared to adevs, can also be seen in the parallel simulation benchmarks.
In the parallel simulation benchmarks, the benefit of our different synchronization protocols is also indicated.

For all benchmarks, results were all well within 1\% deviation of the average, such that only the average is used in the remainder of this section.
The same compilation flags were used for both adevs and dxex benchmarks (``\texttt{-O3 -flto}'').
To guarantee comparable results, no IO was performed during the benchmarking phase.
Simulation traces were used to verify that both adevs and dxex models have exactly the same behaviour.
All benchmarks were performed using Linux, but our simulation tool works equally well on Windows and Mac.

\subsection{Benchmarks}
We use a selection of benchmarks, based on those found in the literature.
Three different types of benchmark are defined, each for a different purpose:
\begin{enumerate}
    \item \textit{Queue} model, based on the HI model of DEVStone~\cite{DEVStone}, creates a chain of atomic models, which are hierarchically nested in each other.
          A single generator will push events into the queue, which get processed by the processors after a fixed or random delay.
          It takes two parameters: width and depth, which determine the widht and depth of the hierarchy.
          This benchmark shows how the complexity of the simulation kernel behaves with an increasing amount of atomic models, and an increasingly deep hierarchy.
          If the processing delay is fixed for all processors, further insight is provided in the collision handling performance of the simulation kernel.
          An example for width 2 and depth 3 is shown in Figure~\ref{fig:queue_model}.

    \item \textit{PHOLD} model, presented by~\cite{PHOLD}.
          It creates a set of $n$ atomic models, and each model has exactly $n-1$ output ports: one for every other atomic model.
          Couplings are made such that each atomic model is directly connected to each other atomic model, such that every atomic model can directly send an event to every other atomic model.
          After a random delay, atomic models will send out an event to a randomly selected output port.
          Output port selection happens in two phases: first it is decided whether the event should be sent to an atomic model inside or outside of this coupled model.
          Afterwards, a uniform selection is made between the possible ports.
          It takes one parameter: the percentage of remote events, which influences the fraction of messages routed to other coupled models.
          This benchmark shows how the simulation kernel behaves in the presence of many local or remote events.
          An example for four models, split over two nodes, is shown in Figure~\ref{fig:PHOLD_model}.

    \item \textit{HighInterconnect} model, a merge of the HI model of DEVStone~\cite{DEVStone} and PHOLD~\cite{PHOLD}.
          It creates a structure similar to the one from PHOLD, but instead of $n-1$ output ports, every atomic model has only a single output port.
          All models are still connected to each other, but through the use of broadcasting: every model will receive a generated event.
          It takes one parameter: the number of models.
          This benchmark investigates the complexity of the routing algorithm.
          An example for four models is shown in Figure~\ref{fig:interconnect_model}.
\end{enumerate}

We opted to deviate from the DEVStone benchmark, as DEVStone tends towards unrealistic models since all internal and external transitions occur simultaneously.
In our benchmark models, there is always the option for simultaneous transition functions (\textit{fixed time advance}), or scattered transition functions (\textit{random time advance}).
Furthermore, they defined the use of an artificial load function, which easily skews the result, making the actual simulation algorithm barely comparable.

@@

\subsection{Sequential Simulation Execution Time}
Despite our core contribution being mainly in the parallel simulation, we still value a comprehensive comparison of sequential simulation results.
First, and foremost, as parallel simulation results are tightly linked to the sequential simulation results: parallel simulation merely adds a synchronization layer over different, essentially sequential, simulation kernels.
Second, since parallel simulation results are validated through the use of adevs.
To provide a more comprehensive comparison to adevs in the parallel simulation benchmarks, sequential simulation results need to be compared.
Only the Queue and HighInterconnect models are relevant for sequential simulation.

\subsubsection{Queue}
In the Queue model, we increase both the width and depth simultaneously, causing a quadratic growth in the number of atomic models.
As can be seen in Figure~\ref{fig:Queue_benchmark}, dxex considerably outperforms adevs.
Through careful analysis of profiling results, we determined that adevs spends much time in handling simulation messages, whereas this is mostly avoided due to the differently designed simulation algorithm of dxex.
Both simulation tools have quadratically increasing execution times, though dxex is much faster thanks to its more efficient simulation control algorithms.

@@

\subsubsection{HighInterconnect}
In the HighInterconnect model, we increase the number of atomic models, thus quadratically increasing the number of couplings.
As can be seen in Figure~\ref{fig:Interconnect_benchmark}, adevs now outperforms dxex by a fair margin.
Analysis showed that this slowdown is caused by the high amount of exchanged events.
Event creation is found to be much slower in dxex than it is in adevs, even despite the use of memory pools in dxex.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{fig/queue_parallel.eps}
	\caption{Queue benchmark results for parallel simulation using variable amount of cores.}
	\label{fig:queue_benchmark_parallel}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{fig/phold_remotes.eps}
    \caption{Phold benchmark results for parallel simulation using 6 cores, with varying fraction of remote events.}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{fig/phold_voorlopig.eps}
    \caption{Phold benchmark results for parallel simulation using 6 cores, with varying amount of high-priority events.}
    \label{fig:phold_priority}
\end{figure}
